Monte Carlo tree search compared to A\(^*\) in airspace configuration decision problem

### 2018-05-20 article/article.tex
(+) More detail on degrouping
(+) Added data

\subsubsection*{Tree translation}
In the tree used in the algorithm exposed earlier, each node has an embedded
state. Considering a node \(u\), its children are the elements of the set of
nodes having as states the possible transitions from the state of \(u\). Going
down in the tree by selecting nodes is then equivalent to moving forward through
time, each step in the tree being a time step.

@@

This expression of the workload, expressed in terms of probabilities, allows us to
use a more complex model for the workload such as a neural network
(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_\text{low}$,
$p_\text{normal}$ and $p_\text{high}$ based on parameters including traffic complexity and
the complexity of the sector (e.g.\ airspace volume).

@@

Considering paragraph~\ref{ssec:theoretical_performances}, the MCTS should
eventually converge to the optimal value. This result reveals a flaw in the
implementation.

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
      legend entries={A\(^*\), MCTS}]
      \foreach \i in {0,1}{%
        \addplot+ table [
          x expr={\lineno}, y index=\i
          ]
          {./data/grouping.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Grouping profile}\label{fig:grouping}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/longterm.data};
      \addplot+[dashed, domain=0:10] {1268};
    \end{axis}
  \end{tikzpicture}
  \caption{The continuous line shows the total cost of a path computed by the
    MCTS algorithm. The x-axis shows the time allowed to the algorithm. The
    dashed curve is 15\% of the optimal cost found with the \(A^*\)
  algorithm.}\label{fig:longterm}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
      \addplot+[dashed, domain=100:800] {18};
    \end{axis}
  \end{tikzpicture}
  \caption{The nice graph}\label{fig:timedepth}
\end{figure}

@@

\subsection{Further work}
Since all the tests has been run using Paris West area, the influence of the
complexity of the airspace has not been assessed. It would be thus relevant to
carry further tests on bigger areas. In~\cite{ferrari2017}, the time needed to
compute the cost with A\(^*\) depends heavily on the traffic itself. Therefore
the use, in this paper, of randomly generated scenarii of traffic may not be
suitable in the context of ATM\@. One can even use a more realistic workload
model such as a neural network as mentioned in~\cite{gianazza2010forecasting}.

>>>>

@@

This expression of the workload, expressed in terms of probabilities, allows us to
use a more complex model for the workload such as a neural network
(see~\cite{gianazza2010forecasting}). This method learns the probabilities $p_\text{low}$,
$p_\text{normal}$ and $p_\text{high}$ based on parameters including traffic complexity and
the complexity of the sector (e.g. airspace volume).

@@

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=timestep, ylabel=num of sectors,
      legend entries={A\(^*\), MCTS}]
      \foreach \i in {0,1}{%
        \addplot+ table [
          x expr={\lineno}, y index=\i
          ]
          {./data/grouping.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Grouping profile}\label{fig:grouping}
\end{figure}
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=time(s), ylabel=cost]
      \addplot+ table {data/shortterm.data};
    \end{axis}
  \end{tikzpicture}
  \caption{Monte Carlo tree search called with varying time allowed to compute
  each step, with a depth of 40 and \(\theta = 11\). The cost seems to be
  approximately stable from \(t=0.5\)s.}\label{fig:shortterm}
\end{figure}

@@

@@

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
      [line width=0.05, mark size=0.1, xlabel=depth, ylabel=cost]
      \foreach \i in {1, 2} {%
        \addplot+ table [ x index = 0, y index = \i ] {./data/cmpdepth.data};
      }
    \end{axis}
  \end{tikzpicture}
  \caption{Cost difference for varying depth with 0.5 seconds allowed to MCTS,
  \(\theta=3\)}\caption{fig:cmpdepth}
\end{figure}

\subsection{Further work}
Since all the tests has been run using Paris West area, the influence of the
complexity of the airspace has not been assessed. It would be thus relevant to
carry further tests on bigger areas. In~\cite{ferrari2017}, the time needed to
compute the cost with A\(^*\) depends heavily on the traffic itself. Therefore
the use, in this paper, of randomly generated scenarii of traffic may not be
suitable in the context of ATM. One can even use a more realistic workload model
such as a neural network as mentioned in~\cite{gianazza2010forecasting}.
